---
title: "Figure 2"
author: "Can Ulutekin"
date: "2025-08-12"
---

### Cleanup

```{r}
rm(list = ls())
```

### Packages

```{r}
library(Seurat)
library(tidyverse)
library(openxlsx)
library(viridis)
```

### Load Data

pns_atlas_sub.rds is the full sural_seurat/pns_atlas.qs2 object from Heming et al but was subset to include only the conditions: VN, CIDP, OIN, CIAP, DPN, ONIN. VN, CIDP and OIN where classified as Inflammed and CIAP, DPN and ONIN where classified as Uninflammed.

Cluster Abbreviations.xlsx is just an excel sheet that holds the general names of very specific clusters and where used to group clusters from the same cell type together.

```{r}
seu_obj_sub <- readRDS(file = "../data/Heming et al/pns_atlas_sub.rds")
clust_abbv <- read.xlsx(xlsxFile = "../data/Heming et al/Cluster Abbreviations.xlsx")
```

### Merge Clusters With Same Cell Types

```{r}
meta_data <- seu_obj_sub@meta.data
clust_order <- seu_obj_sub@misc$cluster_order
clust_col <- seu_obj_sub@misc$cluster_col
active_ident <- seu_obj_sub@active.ident

clust_order <- unique(str_remove(clust_order, "[0-9]"))

clust_col_to_remove <- c("periC2", "periC3", "PC2", "ven_capEC2", "Macro2")
clust_col <- clust_col[!names(clust_col) %in% clust_col_to_remove]

names(clust_col) <- str_remove(names(clust_col), "[0-9]")

meta_data <- meta_data %>%
  dplyr::mutate(cluster = str_remove(cluster, "[0-9]")) %>%
  dplyr::mutate(cluster = factor(cluster, levels = clust_order))

for (i in 1:length(clust_order)) {
  clust_order[i] <- clust_abbv %>%
    dplyr::filter(Abreviation == clust_order[i]) %>%
    .$General.Name
}
clust_order <- unique(clust_order)

clust_col_to_remove <- c("mySC", "repairSC", "damageSC", "endoC", "epiC", "venEC", "artEC", "LEC")
clust_col <- clust_col[!names(clust_col) %in% clust_col_to_remove]

names(clust_col) <- clust_order

rownames_temp <- rownames(meta_data)

meta_data <- meta_data %>%
  dplyr::left_join(clust_abbv %>%
                     dplyr::rename(cluster = Abreviation) %>%
                     dplyr::select(cluster, General.Name),
                   by = "cluster") %>%
  dplyr::select(-cluster) %>%
  dplyr::rename(cluster = General.Name) %>%
  dplyr::relocate(cluster, .before = "level2") %>%
  dplyr::mutate(cluster = factor(cluster, levels = clust_order))

rownames(meta_data) <- rownames_temp

seu_obj_sub@meta.data <- meta_data
seu_obj_sub@misc$cluster_order <- clust_order
seu_obj_sub@misc$cluster_col <- clust_col

Idents(seu_obj_sub) <- seu_obj_sub@meta.data$cluster

rm(meta_data, clust_order, clust_col, clust_col_to_remove, rownames_temp, i) # cleanup
```

### Figure 2A

```{r}
p <- DimPlot(
  seu_obj_sub,
  reduction = "umap.scvi.full",
  pt.size = 0.1,
  alpha = 0.1,
  cols = seu_obj_sub@misc$cluster_col,
  label = TRUE,
  raster = FALSE
) +
  NoLegend() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) +
  xlab("UMAP1") +
  ylab("UMAP2") +
  coord_fixed()

pdf("../figures/Figure2A.pdf")
print(p)
invisible(dev.off())

png("../figures/Figure2A.png")
print(p)
invisible(dev.off())

rm(p) # cleanup
```

### Figure 2B

```{r}
p <- Seurat::DotPlot(seu_obj_sub,
                features = c("CXCL8", "CXCR1", "CXCR2"),
                scale = TRUE) +
  viridis::scale_color_viridis(option = "viridis") +
  coord_flip() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank()
  )

pdf("../figures/Figure2B.pdf", width = 9, height = 4)
print(p)
invisible(dev.off())

rm(p) # cleanup
```

### Figure 2C

```{r}
p <- Seurat::DotPlot(seu_obj_sub,
                features = c("LIF", "LIFR", "IL6ST"),
                scale = TRUE) +
  viridis::scale_color_viridis(option = "viridis") +
  coord_flip() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank()
  )

pdf("../figures/Figure2C.pdf", width = 9, height = 4)
print(p)
invisible(dev.off())

rm(p) # cleanup
```
